---
title: "Using Mordecai Geoparsing In R"
author: "Andy Halterman"
output: html_document
references:
- id: hrtexts
  title: "Human Rights Texts: Converting Human Rights Primary Source Documents into Data"
  author:
  - family: Fariss
    given: Christopher J.
  - family: Linder
    given: Fridolin J.
  - family: Crabtree
    given: Charles D.
  - family: Biek
    given: Megan A.
  - family: Ross
    given: Ana-Sophia M.
  - family: Kaur
    given: Taranamol
  - family: Tsai
    given: Michael
  year: 2015
  DOI: 10.7910/DVN/IAH8OY
  url : "http://dx.doi.org/10.7910/DVN/IAH8OY"
  publisher: Harvard Dataverse
---

One of the advantages of Mordecai's HTTP-based interface is that any language
that can make HTTP POST requests can interact with it without needed special
Mordecai packages or code. This example demonstrates how to read in a text file,
have Mordecai geolocate it to the country level, and then do a full geoparse
with Mordecai. It then shows how to format the returned data and easily plot it
on a map.

For this demonstration, we need `httr` for handing the request to Mordecai,
`dplyr` for formatting the result, and `leaflet` for making a quick interactive
map of the results.

```{r message = FALSE, warning=FALSE}
library(httr)
library(dplyr)
library(leaflet)
```

Set the endpoints for the `country` and `places` endpoints. Here, Mordecai is
running locally.

```{r}
country_url <- "http://localhost:5000/country"
places_url <- "http://localhost:5000/places"
```

We can then make a GET request to Mordecai to make sure it's up and running and
that we can talk to it.

```{r}
t <- GET(url = country_url, as = "parsed")
content(t)
```

This response lets us know that it is and gives us some guidance on what data
format it expects.

First, let's test Mordecai's country coding capability. We can read in one of
the human rights texts prepared by @hrtexts...

```{r}
bol <- paste(readLines("BOL_2009_Amnesty_International.txt"), collapse = " ")
```

...and then POST it to the `country` endpoint.

```{r}
bol_country <- POST(url = country_url, 
                   as = "parsed", 
                   body = list("text" = bol), 
                   encode = "json")

content(bol_country)
```

Thankfully, since this is indeed a text about Bolivia, Mordecai codes it as `BOL`. 

Now let's do a full geoparsing, extracting all the place names in the text and 
finding their correct entries in the gazetteer. The final line formats the
response as a dataframe.

```{r}
bol_places <- POST(url = places_url, 
                   as = "parsed", 
                   body = list("text" = bol), 
                   encode = "json")

bol_places_df <- bind_rows(content(bol_places))
bol_places_df
```

These locations pass an eyeball test: no placename was located to a completely 
different looking place. Now, for fun, we can plot these locations on an
interactive leaflet map, sized according to their mentions in the text.

```{r message = FALSE}
bol_places_df %>% 
  group_by(placename) %>% 
  mutate(count = n()) %>% 
  distinct() %>% 
  leaflet(.) %>% 
    addTiles() %>%
    addCircleMarkers(popup = ~placename, radius = ~3*(count + 2))
```

A more serious example will use many more texts than this one and would probably
wrap the raw POST requests into a function. But hopefully this example will get
R users started with Mordecai.

# References
